---
title: "Individual Parameter Fitting"
author: "Alana"
date: "2022/5/23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,warning = FALSE, message = FALSE)
```

# Parameter fitting
To start the parameter estimation for fitting spiking behaviour of model to oxytocin neurons, we first tried to adjust several key parameters alone and explored their influences on firing rate and ISI distribution.  
  
## Fit Ire
```{r}
Vrest = -56 # Resting membrane potential
Vthresh = -50 # Threshold of action potential

# Synaptic input
eh = 2 # The EPSP magnitude
ih = -2 # The ISPS magnitude
lambda_syn = 3.5 # The PSP decay rate
Ires=c(seq(260,270,2),seq(460,470,2),seq(290,300,2),seq(580,590,2)) # The input rate (Hz)
Iratio = 1 # The ratio of IPSPs to EPSPs

# Post-spike potential
kHAP = 30 # Magnitude parameter of hyperpolarising afterpotential (HAP)
lambda_HAP = 7.5 # Half-life parameter of HAP
kAHP = 0.2 # Magnitude parameter of after hyperpolarising potential (AHP)
lambda_AHP = 350 # Half-life parameter of AHP

# Half-life parameter are converted to time strain
tausyn = log(2) / lambda_syn 
tauHAP = log(2) / lambda_HAP
tauAHP = log(2) / lambda_AHP
Ire_fit=data.frame()

# Simulation time
runtime = 1000 # simulation time in second
numsteps = runtime * 1000 # simulation time in ms

for(Ire in Ires){
  # Initialize values
  Vsyn = 0 # Synaptic input signal
  HAP = 0 
  AHP = 0
  time = 0
  hstep = 1 #(the time step in ms) 
  spikecount = 0
  tepsp = 0
  tipsp = 0
  epsprate = Ire / 1000 
  ipsprate = epsprate * Iratio
  spiketime=c()
  
  for(i in 1:numsteps) { 
    # generate EPSPs
    nepsp = 0
    while (tepsp < hstep) {
      rand = runif(1)
      nepsp = nepsp + 1
      tepsp = - log(1 - rand)/epsprate+tepsp
    }
    tepsp = tepsp - hstep 
    # generate IPSPs 
    nipsp = 0
    while (tipsp < hstep) {
      rand = runif(1)
      nipsp = nipsp + 1
      tipsp = - log(1 - rand) / ipsprate + tipsp
    }
    tipsp = tipsp - hstep
    inputsyn = nepsp * eh + nipsp * ih
    
    # membrane variables
    Vsyn = Vsyn + inputsyn - Vsyn * tausyn 
    HAP = HAP - HAP * tauHAP
    AHP = AHP - AHP * tauAHP
    V = Vrest + Vsyn - HAP - AHP
    
    if(V > Vthresh) {
    spikecount = spikecount + 1
    spiketime[spikecount] = i
  
    # add the post-spike potentials
    HAP = HAP + kHAP
    AHP = AHP + kAHP
    }}
  spiketime=spiketime/1000 # Convert to (s) scale
  ds=cbind(rep(Ire,length(spiketime)),spiketime)
  Ire_fit=rbind(Ire_fit,ds)
}
colnames(Ire_fit)[1]='Ire'
```

### Calculate spiking rate 

```{r}
spike_rate=c()
for(Ire in Ires){
  spiketimedt=Ire_fit[which(Ire_fit$Ire==Ire),]
  spike_rate=c(spike_rate,length(spiketimedt$Ire)/max(spiketimedt$spiketime))
}
library(knitr)
kable(data.frame(cbind(Ires,spike_rate)))
```
  
The firing rates of four neurons are 3.39, 7.27, 3.99 and 9.24 spike/s.According to the simulation, the approximated input rate Ire for each neuron is 267, 460, 294 and 585.   
  
### Calculate  ISI
```{r}
Ire_fit_ISI=data.frame()
for(Ire in Ires){
  spiketimedt=Ire_fit[which(Ire_fit$Ire==Ire),]
  ISI=lapply(1:(length(spiketimedt$spiketime)-1),
             function(i){spiketimedt$spiketime[i+1]-spiketimedt$spiketime[i]})
  ISI=as.vector(unlist(ISI))
  ds=cbind(rep(Ire,length(ISI)),ISI)
  Ire_fit_ISI=rbind(Ire_fit_ISI,ds)
}
Ire_fit_ISI=data.frame(Ire_fit_ISI)
colnames(Ire_fit_ISI)[1]='Ire'
```
### Plot ISI
```{r}
library(ggplot2)
library(ggpubr)
Ire_fit_ISI$Ire=as.factor(Ire_fit_ISI$Ire)
plot_Ire_fit_ISI=ggplot(Ire_fit_ISI,aes(x=ISI,color=Ire))+
  geom_density(binwidth = 0.005)+
  xlim(-0.1,0.5)+ylim(0,12)+
  xlab('ISI (s)')+ylab('Count')+
  theme_classic()
ggplot(Ire_fit_ISI,aes(x=ISI,color=Ire))+
  geom_histogram(binwidth = 0.005,aes(fill=Ire),alpha=0.5)+
  xlim(0,0.5)+xlab('ISI (s)')+ylab('Count')+
  theme_classic()
ggsave("/Users/chenghui/Documents/synoptic/cell data/Ire_fit_ISI.png",plot_Ire_fit_ISI)
```

## Fit lambda_HAP
```{r}
Vrest = -56 # Resting membrane potential
Vthresh = -50 # Threshold of action potential

# Synaptic input
eh = 2 # The EPSP magnitude
ih = -2 # The ISPS magnitude
lambda_syn = 3.5 # The PSP decay rate
Ire=300 # The input rate /s
Iratio = 1 # The ratio of IPSPs to EPSPs

# Post-spike potential
kHAP = 30 # Magnitude parameter of hyperpolarising afterpotential (HAP)
lambda_HAPs = seq(4,12,2) # Half-life parameter of HAP
kAHP = 0.2 # Magnitude parameter of after hyperpolarising potential (AHP)
lambda_AHP = 350 # Half-life parameter of AHP

# Half-life parameter are converted to time strain
tausyn = log(2) / lambda_syn 
#tauHAP = log(2) / lambda_HAP
tauAHP = log(2) / lambda_AHP

#run time
runtime = 1000 # simulation time in second
numsteps = runtime * 1000 # simulation time in ms
```

```{r}
lambda_HAP_fit=data.frame()
for(lambda_HAP in lambda_HAPs){
  #print(lambda_HAP)
  # Initialize values
  Vsyn = 0 # Synaptic input signal
  HAP = 0 
  AHP = 0
  time = 0
  hstep = 1 #(the time step in ms) 
  spikecount = 0
  tepsp = 0
  tipsp = 0
  tauHAP = log(2) / lambda_HAP
  spiketime=c()
  
  for(i in 1:numsteps) { 
    # generate EPSPs
    nepsp = 0
    while (tepsp < hstep) {
      rand = runif(1)
      nepsp = nepsp + 1
      tepsp = - log(1 - rand)/epsprate+tepsp
    }
    tepsp = tepsp - hstep 
    #// generate IPSPs 
    nipsp = 0
    while (tipsp < hstep) {
      rand = runif(1)
      nipsp = nipsp + 1
      tipsp = - log(1 - rand) / ipsprate + tipsp
    }
    tipsp = tipsp - hstep
    inputsyn = nepsp * eh + nipsp * ih
 
    # membrane variables
    Vsyn = Vsyn + inputsyn - Vsyn * tausyn 
    HAP = HAP - HAP * tauHAP
    AHP = AHP - AHP * tauAHP
    V = Vrest + Vsyn - HAP - AHP
    
    if(V > Vthresh) {
    spikecount = spikecount + 1
    spiketime[spikecount] = i
  
    # add the post-spike potentials
    HAP = HAP + kHAP
    AHP = AHP + kAHP
    }}
  spiketime=spiketime/1000 # Convert to (s) scale
  ds=cbind(rep(paste0('lambda_HAP_',lambda_HAP),length(spiketime)),spiketime)
  lambda_HAP_fit=rbind(lambda_HAP_fit,ds)
}
colnames(lambda_HAP_fit)[1]='lambda_HAP'
```
### Calculate ISI
```{r}
lambda_HAP_fit_ISI=data.frame()
for(lambda_HAP in lambda_HAPs){
  spiketimedt=lambda_HAP_fit[which(lambda_HAP_fit$lambda_HAP==paste0('lambda_HAP_',lambda_HAP)),]
  spiketimedt$spiketime=as.numeric(as.character(spiketimedt$spiketime))
  ISI=c()
  for (i in 1:(length(spiketimedt$spiketime)-1)){
    ISI=c(ISI,spiketimedt$spiketime[i+1]-spiketimedt$spiketime[i])
  }
  ds=cbind(rep(paste0('lambda_HAP_',lambda_HAP),length(ISI)),ISI)
  lambda_HAP_fit_ISI=rbind(lambda_HAP_fit_ISI,ds)
}
lambda_HAP_fit_ISI=data.frame(lambda_HAP_fit_ISI)
colnames(lambda_HAP_fit_ISI)[1]='Cell_ID'
lambda_HAP_fit_ISI$ISI=as.numeric(as.character(lambda_HAP_fit_ISI$ISI))
```
### Plot ISI
```{r}
lambda_HAP_fit_ISI$Cell_ID=as.factor(lambda_HAP_fit_ISI$Cell_ID)
plot_lambda_HAP_fit_ISI=ggplot(lambda_HAP_fit_ISI,aes(x=ISI,color=Cell_ID))+
  geom_histogram(binwidth = 0.005,aes(fill=Cell_ID),alpha=0.5)+
  xlim(-0,0.5)+theme_classic()+xlab('ISI (s)')+ylab('Count')
ggplot(lambda_HAP_fit_ISI,aes(x=ISI,color=Cell_ID))+
  geom_density()+xlim(0,0.5)+theme_classic()+
  xlab('ISI (s)')+ylab('Count')
ggsave("/Users/chenghui/Documents/synoptic/cell data/plot_lambda_HAP_fit_ISI.png",
       plot_lambda_HAP_fit_ISI)
# Ire_fit_ISI_300=Ire_fit_ISI[which(Ire_fit_ISI$Ire==300),]
# ggplot(Ire_fit_ISI_300,aes(x=ISI,color=Ire))+
# geom_density(binwidth = 0.001)+xlim(-0.5,0.5)+ylim(0,1000)+
# theme_classic()+xlab('ISI (s)')+ylab('Count')
```
## Fit kHAP
### Parameter Initialization
```{r}
Vrest = -56 # Resting membrane potential
Vthresh = -50 # Threshold of action potential

# Synaptic input
eh = 2 # The EPSP magnitude
ih = -2 # The ISPS magnitude
lambda_syn = 3.5 # The PSP decay rate
Ire=300 # The input rate /s
Iratio = 1 # The ratio of IPSPs to EPSPs
epsprate = Ire / 1000 
ipsprate = epsprate * Iratio


# Post-spike potential
kHAPs = seq(20,100,20) # Magnitude parameter of hyperpolarising afterpotential (HAP)
lambda_HAP = 7.5 # Half-life parameter of HAP
kAHP = 0.2 # Magnitude parameter of after hyperpolarising potential (AHP)
lambda_AHP = 350 # Half-life parameter of AHP

# Half-life parameter are converted to time strain
tausyn = log(2) / lambda_syn 
#tauHAP = log(2) / lambda_HAP
tauAHP = log(2) / lambda_AHP

#run time
runtime = 1000 # simulation time in second
numsteps = runtime * 1000 # simulation time in ms
```

### Model
```{r}
kHAP_fit=data.frame()
for(kHAP in kHAPs){
  #print(kHAP)
  # Initialize values
  Vsyn = 0 # Synaptic input signal
  HAP = 0 
  AHP = 0
  time = 0
  hstep = 1 #(the time step in ms) 
  spikecount = 0
  tepsp = 0
  tipsp = 0
  tauHAP = log(2) / lambda_HAP
  spiketime=c()
  
  for(i in 1:numsteps) { 
    # generate EPSPs
    nepsp = 0
    while (tepsp < hstep) {
      rand = runif(1)
      nepsp = nepsp + 1
      tepsp = - log(1 - rand)/epsprate+tepsp
    }
    tepsp = tepsp - hstep 
    #// generate IPSPs 
    nipsp = 0
    while (tipsp < hstep) {
      rand = runif(1)
      nipsp = nipsp + 1
      tipsp = - log(1 - rand) / ipsprate + tipsp
    }
    tipsp = tipsp - hstep
    inputsyn = nepsp * eh + nipsp * ih
 
    # membrane variables
    Vsyn = Vsyn + inputsyn - Vsyn * tausyn 
    HAP = HAP - HAP * tauHAP
    AHP = AHP - AHP * tauAHP
    V = Vrest + Vsyn - HAP - AHP
    
    if(V > Vthresh) {
    spikecount = spikecount + 1
    spiketime[spikecount] = i
  
    # add the post-spike potentials
    HAP = HAP + kHAP
    AHP = AHP + kAHP
    }}
  spiketime=spiketime/1000 # Convert to (s) scale
  ds=cbind(rep(paste0('kHAP_',kHAP),length(spiketime)),spiketime)
  kHAP_fit=rbind(kHAP_fit,ds)
}
colnames(kHAP_fit)[1]='kHAP'
```
### Calculate ISI
```{r}
kHAP_fit_ISI=data.frame()
for(kHAP in kHAPs){
  spiketimedt=kHAP_fit[which(kHAP_fit$kHAP==paste0('kHAP_',kHAP)),]
  spiketimedt$spiketime=as.numeric(as.character(spiketimedt$spiketime))
  ISI=c()
  for (i in 1:(length(spiketimedt$spiketime)-1)){
    ISI=c(ISI,spiketimedt$spiketime[i+1]-spiketimedt$spiketime[i])
  }
  ds=cbind(rep(paste0('kHAP_',kHAP),length(ISI)),ISI)
  kHAP_fit_ISI=rbind(kHAP_fit_ISI,ds)
}
kHAP_fit_ISI=data.frame(kHAP_fit_ISI)
colnames(kHAP_fit_ISI)[1]='Cell_ID'
kHAP_fit_ISI$ISI=as.numeric(as.character(kHAP_fit_ISI$ISI))
```
### Plot ISI
```{r}
kHAP_fit_ISI$Cell_ID=as.factor(kHAP_fit_ISI$Cell_ID)
plot_kHAP_fit_ISI=ggplot(kHAP_fit_ISI,aes(x=ISI,color=Cell_ID))+
  geom_histogram(binwidth = 0.005,aes(fill=Cell_ID),alpha=0.5)+xlim(0,0.8)+
  theme_classic()+xlab('ISI (s)')+ylab('Count')
plot_kHAP_fit_ISI=ggplot(kHAP_fit_ISI,aes(x=ISI,color=Cell_ID))+
  geom_density(binwidth = 0.005)+xlim(0,0.8)+
  theme_classic()+xlab('ISI (s)')+ylab('Count')
#ggsave("/Users/chenghui/Documents/synoptic/cell data/plot_kHAP_fit_ISI.png",plot_kHAP_fit_ISI)
# Ire_fit_ISI_300=Ire_fit_ISI[which(Ire_fit_ISI$Ire==300),]
# ggplot(Ire_fit_ISI_300,aes(x=ISI,color=Ire))+
# geom_density(binwidth = 0.001)+xlim(-0.5,0.5)+ylim(0,1000)+
# theme_classic()+xlab('ISI (s)')+ylab('Count')
```

# Compare with cell data
## Read in data
```{r}
home="/Users/chenghui/Documents/synoptic/cell data/"
Cell_ISI=read.csv(paste0(home,"Cell_ISI.csv"))
Cell_Hazard=read.csv(paste0(home,"Cell_Hazard.csv"))
Cell_ID=unique(Cell_ISI$Cell_ID)
ID=1
```

## lambda_HAP estimation
```{r}
ISI_combine=rbind(Cell_ISI[which(Cell_ISI$Cell_ID==Cell_ID[ID]),2:3],lambda_HAP_fit_ISI)
ggplot(ISI_combine,aes(x=ISI,color=Cell_ID))+
  geom_histogram(aes(fill=Cell_ID),alpha=0.5,binwidth = 0.005)+xlim(-0.1,1)+
  theme_classic()+xlab('ISI (s)')+ylab('Count')
ggplot(ISI_combine,aes(x=ISI,color=Cell_ID))+geom_density()+
  xlim(-0.1,1)+theme_classic()+xlab('ISI (s)')+ylab('Count')
```

## kHAP estimation
```{r}
ISI_combine=rbind(Cell_ISI[which(Cell_ISI$Cell_ID==Cell_ID[ID]),2:3],kHAP_fit_ISI)
ggplot(ISI_combine,aes(x=ISI,color=Cell_ID))+
  geom_histogram(binwidth = 0.005)+xlim(-0.1,1)+
  theme_classic()+xlab('ISI (s)')+ylab('Count')
ggplot(ISI_combine,aes(x=ISI,color=Cell_ID))+geom_density()+
  xlim(-0.1,1)+theme_classic()+xlab('ISI (s)')+ylab('Count')
```
