---
title: "Analyisis of in vivo Spike Timing of Oxytocin Neurons"
author: "Alana"
date: "2022/5/18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,warning = FALSE, message = FALSE)
```

# Read in Data
The spiking data were provided by Dr. Duncan Macgregor as txt format and read into R.
```{r}
home="/Users/chenghui/Documents/synoptic/cell data/"
data= list()
cellnum=0
Cell_ID=c()
for (i in list.files(home)){
  if (length(grep('txt', i))==0){next}
   cellnum=cellnum+1
  # get file prefix
  prefix = strsplit(i, split='\\.')[[1]][1]
  Cell_ID=c(Cell_ID,prefix)
  dt = read.table(file.path(home, i),header = T)
  data[[cellnum]]=dt
}
```

# Spike Rate
Calculate the frequency of firing spikes.
```{r}
Frequency=c()
for (i in 1:cellnum) {
  celldt=data[[i]]
  Frequency=c(Frequency,length(celldt[,1])/celldt[length(celldt[,1]),1])
}
Frequency_dt=data.frame(Cell_ID,Frequency)
colnames(Frequency_dt)[2]='Frequency (spike/s)'
print(Frequency_dt)
library(knitr)
kable(Frequency_dt)
```
# Spike Histogram
Plot the distribution for spikes across time.
```{r}
library(ggplot2)
library(ggpubr)
plotspike=list()
for (i in 1:cellnum){
  celldt=data[[i]]
  colnames(celldt)='spike_time'
  #plotspike[[i]]=hist(celldt[,1],col="skyblue",breaks=seq(0,max(celldt[,1]+10,10)),main=Cell_ID[i])
  plotspike[[i]]=ggplot(celldt,aes(x=spike_time))+
    geom_histogram(binwidth = 10,fill='grey')+
    theme_classic()+xlab('Spike Time (s)')+
    labs(title = Cell_ID[i])
}

ggsave(paste0(home,"spike.png"),plot=ggarrange(plotlist = plotspike))
```

# Inter-Spike Interval (ISI) Analysis
Calculate ISIs based on spike timing data.
```{r}
ISI=list()
plot=list() # save histogram plottings
plothist=list()
for(i in 1:cellnum){
  celldt=data[[i]]
  isi=c()
  for(j in 1:(length(celldt[,1])-1)){
    isi=c(isi,celldt[j+1,1]-celldt[j,1])
  }
  plothist[[i]]=hist(isi,breaks=seq(0,max(isi)+0.005,0.005),plot = F)
  isi=data.frame(isi)
  plot[[i]]=ggplot(isi,aes(x=isi))+
    geom_histogram(aes(x=isi),binwidth=0.005,color='grey',fill='grey')+
    theme_classic()+labs(title = Cell_ID[[i]])+xlab('ISI (s)')
  ISI[[i]]=isi
}

ggsave(paste0(home,"ISI.png"),plot=ggarrange(plotlist = plot))

# Save ISI data as data frame for consecutive parameter fitting
ISIdt=data.frame()
for (i in 1:cellnum){
  dt=cbind(rep(Cell_ID[i],length(ISI[[i]])),ISI[[i]])
  ISIdt=rbind(ISIdt,dt)
}
colnames(ISIdt)=c('Cell_ID','ISI')
write.csv(ISIdt,paste0(home,"Cell_ISI.csv"))
```
#ISI count normalization
Normalised ISI count to percentage, for a better comparison among neurons.
```{r}
plotnormhist=list()
for (i in 1:cellnum) {
  ISIhist=plothist[[i]]
  ISIhist$counts=ISIhist$counts/sum(ISIhist$counts)
  #plot(ISIhist)
  #axis(side=1,c(0,3,0.5))
  #axis(side=2,c(0,800,200))
  ISInorm=data.frame(ISIhist$breaks[1:length(ISIhist$breaks)-1]+
                       (ISIhist$breaks[2]-ISIhist$breaks[1])/2,ISIhist$counts)
  colnames(ISInorm)=c('ISI','norm_count')
  plotnormhist[[i]]=ggplot(ISInorm,aes(x=ISI,y=norm_count))+
    geom_bar(stat='identity',width = 0.005,fill='grey')+
    theme_classic()+scale_y_continuous(expand = c(0,0),limits = c(0,0.05))+
    xlab('ISI (s)')+ylab('Count (%)')+xlim(0,0.5)+labs(title=Cell_ID[[i]])
}
ggsave(paste0(home,"NormISI.png"),plot=ggarrange(plotlist = plotnormhist))

```

# Fit Poison
Fit the ISI ditribution to Poison distribution
```{r}
poi=list()
for(i in 1:cellnum){
  ISIhist=plothist[[i]]
  poi[[i]]=glm(ISIhist$counts~ISIhist$breaks[1:length(ISIhist$counts)],family = poisson())
  print(Cell_ID[i])
  print(summary(poi[[i]]))
}
```
# The hazard function
The hazard replaces the absolute probabilities of the histogram with conditional probabilities. It plots the chance of firing a spike (neuron excitability) against the time since the last spike.  

## Parameters used for calculation
*spikecount* – the number of spike times    
*hist* – a vector containing the histogram in 5ms bins    
*haz* – a vector to store the hazard    
*bincount* – the number of bins in the histogram, typically 100       

## Hazard function 
hazard in bin [t, t + 5]) = (number of ISIs in bin [t, t + 5])/(number of ISIs of length > t)    
```{r}
Hazarddt=data.frame()
plothazard=list()
for (i in 1:cellnum){
  celldt=data[[i]]
  spikecount=length(celldt[,1])
  hist=plothist[[i]]$counts
  haz=c()
  bincount=length(plothist[[i]]$counts)
  hazcount=0
  for(j in 1:bincount){
    haz[j]=hist[j]/(spikecount-hazcount)
    hazcount=hazcount+hist[j]
  }
  #plot(c(1:bincount)*0.005,haz)
  time=c(1:bincount)*0.005
  hazarddt=data.frame(time,haz)
  plothazard[[i]]=ggplot(hazarddt,aes(x=time,y=haz))+
    geom_point(color='#1F77B4FF',size=2,alpha=0.6)+
    theme_classic()+labs(title = Cell_ID[i])+xlab('Time (s)')+ylab('Hazard')+
    xlim(0,0.5)+ylim(0,0.2)
  
  # Save Hazard data for parameter fittting
  dt=cbind(rep(Cell_ID[i],nrow(hazarddt)),hazarddt)
  Hazarddt=rbind(Hazarddt,dt)
}
ggsave(paste0(home,"Hazard.png"),plot=ggarrange(plotlist = plothazard))

colnames(Hazarddt)[1]='Cell_ID'
write.csv(Hazarddt,paste0(home,"Cell_Hazard.csv"))
```

# Spike train analysis
```{r}
STA=list()
for (i in 1:cellnum){
  sta=c()
  celldt=data[[i]]
  for (j in 2:nrow(celldt)) {
    if(j<=20){
      sta=c(sta,mean(celldt[1:(j-1),])/celldt[j,])
    }else{
      sta=c(sta,mean(celldt[(j-20):(j-1),])/celldt[j,])
    }
  }
  STA[[i]]=sta
}
```


